Shapes analysis for time series.

### 2024-05-14 NEURIPS24/neurips_2024_to_submit.tex
update

\begin{abstract}
    %TO CHANGE]
    %Analyzing the inter-individual variability within a cluster of time series is particularly appealing in medicine and biology.
    %Individuals have specificities depending on their genes or disease
     %while sharing, on average, a similar pattern in the features of their corresponding time series. For instance, an electrocardiogram presents a typical shape of the heartbeat, which varies depending on the individual.
    %In this paper, we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
    %Inspired by the shape analysis literature, we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) to time series, considering the temporal evolution as their distinctive feature.  
    % Given a dataset of $N$ individual's time series $(s^j)_{j\in[N]}$,
    %  we learn a time series of reference $\mathbf{s}_0$ encoding the common part of the individuals as well as diffeomorphisms $(\phi^j)_{j\in[N]}$ encoding the specificities of each individual:
    %  $s^j$ is seen as the transformation of $\mathbf{s}_0$ by $\phi_j$.
    %  Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
    This paper introduces an unsupervised representation learning (URL) algorithm for time series tailored to biomedical inter-individual studies. The proposed algorithm manages irregularly sampled multivariate time series of variable lengths and provides shape-based representations of temporal data.
    At the crossroads between URL for time series and shape analysis, the algorithm extends the Large Deformation Diffeomorphic Metric Mapping (LDDMM) to the case of time series. This extension is a nontrivial application of the LDDMM framework since it involves applying deformations to the graph of time series, which have more structure than curves due to the inclusion of the temporal dimension. 
    In this work, we establish a representation theorem for the graph of a time series and derive its consequences on the LDDMM framework. We showcase the advantages of our representation compared to existing methods using synthetic data and real-world examples motivated by biomedical applications.
\end{abstract}

>>>>

\begin{abstract}
    %TO CHANGE]
    Analyzing the inter-individual variability within a cluster of time series is particularly appealing in medicine and biology.
    Individuals have specificities depending on their genes or disease
     while sharing, on average, a similar pattern in the features of their corresponding time series. For instance, an electrocardiogram presents a typical shape of the heartbeat, which varies depending on the individual.
    %In this paper, we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
    %Inspired by the shape analysis literature, we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) to time series, considering the temporal evolution as their distinctive feature.  
    % Given a dataset of $N$ individual's time series $(s^j)_{j\in[N]}$,
    %  we learn a time series of reference $\mathbf{s}_0$ encoding the common part of the individuals as well as diffeomorphisms $(\phi^j)_{j\in[N]}$ encoding the specificities of each individual:
    %  $s^j$ is seen as the transformation of $\mathbf{s}_0$ by $\phi_j$.
    %  Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
    This paper introduces an unsupervised representation learning (URL) algorithm for time series tailored to biomedical inter-individual studies.
    The idea is to represent time series as the deformation of a time series of reference using Large Deformation Diffeomorphic Metric Mapping (LDDMM).
     Once the deformations and time series of reference are learned, the vector representations of time series are then given by the parametrization of the deformation.
     At the crossroads between URL for time series and shape analysis, the proposed algorithm manages irregularly sampled multivariate time series of variable lengths and provides shape-based representations of temporal data.
    %At the crossroads between URL for time series and shape analysis. %the algorithm extends the Large Deformation Diffeomorphic Metric Mapping (LDDMM) to the case of time series. This extension is a nontrivial application of the LDDMM framework since it involves applying deformations to the graph of time series, which have more structure than curves due to the inclusion of the temporal dimension. 
    In this work, we establish a representation theorem for the graph of a time series and derive its consequences on the LDDMM framework. We showcase the advantages of our representation compared to existing methods using synthetic data and real-world examples motivated by biomedical applications.
\end{abstract}
